agdarsec -- Total Parser Combinators

### 2017-04-19 doc/tyde17/main.tex
[ doc ] more about box

\newcommand{\shgrab}[1]{\bigskip\ExecuteMetaData[parser.tex]{#1}}
\newcommand{\hgrab}[1]{\ExecuteMetaData[parser.tex]{#1}}

\newcommand{\sagrab}[2]{\bigskip\ExecuteMetaData[#1.tex]{#2}}
\newcommand{\agrab}[2]{\ExecuteMetaData[#1.tex]{#2}}

@@

Finally, the last kind of combinator\todo{cite}
Instead of defining it as a

@@

\sagrab{Induction/Nat/Strong}{extract}
\agrab{Induction/Nat/Strong}{duplicate}

\sagrab{Induction/Nat/Strong}{map}
\agrab{Induction/Nat/Strong}{app}

\agrab{Induction/Nat/Strong}{fix}

\paragraph{Remark} The type of \AF{fix} corresponds exactly to
strong induction for the natural numbers. A similar \AR{□} type
constructor can be defined for any induction principle relying
on an accessibility predicate. Which means that a library's
types can be cleaned up by using these combinators in any
situation where one had to give up structural induction for a
more powerful alternative.

>>>>

\newcommand{\shgrab}[1]{\medskip\ExecuteMetaData[parser.tex]{#1}}
\newcommand{\hgrab}[1]{\ExecuteMetaData[parser.tex]{#1}}

\newcommand{\sagrab}[2]{\medskip\ExecuteMetaData[#1.tex]{#2}}
\newcommand{\agrab}[2]{\ExecuteMetaData[#1.tex]{#2}}
\newcommand{\BOX}{\AR{□}}

@@

We can already see the benefits of using these aliases. For instance
the fairly compact expression \AF{[} (\AB{P} \AF{⊗} \AB{Q}) \AF{⟶} \AB{R} \AF{]}
corresponds to the more verbose type
\AS{∀} \{\AB{n}\} \AS{→} (\AB{P} \AB{n} \AF{×} \AB{Q} \AB{n}) \AS{→} \AB{R} \AB{n}.

Last but not least, the \BOX{} type constructor takes a \AD{ℕ}-indexed
set and produces the one corresponding to the allowed recursive
calls in a function defined by course-of-values recursion: \BOX{}
\AB{A} \AB{n} associates to each \AB{m} strictly smaller than \AB{n}
a value of type \AB{A} \AB{m}.

@@

\paragraph{Remark: Record Wrapper} Instead of defining \BOX{} as a
function like the other combinators, we wrap the corresponding
function space in a record type. This prevents normalisation from
unfolding the combinator too eagerly and makes types more readable
during interactive development.

\paragraph{Remark: Irrelevance} The argument stating that \AB{m}
is strictly smaller than \AB{n} is preceded by a dot. In Agda,
it means that this value is irrelevant and can be erased by the
compiler. In Coq, we would define the relation \AF{\_<\_} in
\type{Prop} to achieve the same.

\medskip{}

This \BOX{} type combinator has some useful value combinators.
The first thing we can notice is the fact that \textbf{\BOX{} is a
functor}; that is to say that given a natural transformation from
\AB{A} to \AB{B}, we can define a natural transformation from
\BOX{} \AB{A} to \BOX{} \AB{B}.

\sagrab{Induction/Nat/Strong}{map}

\paragraph{Remark: Copatterns} The definition of \AF{map} uses
the \BOX{} field named \ARF{call} on the \emph{left hand side}.
This is a copattern\todo{cite}, meaning that we explain how
the definition is \emph{observed} (via \ARF{call}) rather than
\emph{constructed} (via \AIC{mkBox}).

\medskip{}

Because \AF{\_<\_} is defined in terms of \AD{\_≤\_}, \AF{≤-refl}
which is the proof that \AD{\_≤\_} is reflexive is also a proof
that any \AB{n} is strictly smaller than \AN{1} \AF{+} \AB{n}.
We can use this fact to write the following \AF{extract} function:

\sagrab{Induction/Nat/Strong}{extract}

\paragraph{Remark: Counit} The careful reader will have noticed
that this is not quite the extract we would expect from a comonad:
for a counit, we would need a natural transformation between
\BOX{} \AB{A} and \AB{A} i.e. a function of type
\AF{[} \BOX{} \AB{A} \AF{⟶} \AB{A} \AF{]}. We will not be able
to define such a function: \BOX{} \AB{A} \AN{0} is isomorphic to
the unit type so we would have to generate an \AB{A} \AN{0} out
of thin air.

\medskip

Even though we cannot have a counit, we are still able to define
a comultiplication thanks to the fact that \AF{\_<\_} is transitive.

\sagrab{Induction/Nat/Strong}{duplicate}

\agrab{Induction/Nat/Strong}{app}
\sagrab{Induction/Nat/Strong}{fix}

\paragraph{Remark: Generalisation} The type of \AF{fix} corresponds
exactly to strong induction for the natural numbers. A similar \BOX{}
type constructor can be defined for any induction principle relying
on an accessibility predicate. Which means that a library's types
can be cleaned up by using these combinators in any situation where
one had to give up structural induction for a more powerful alternative.
